Statistical analysis script

### 2023-04-07 code/analysis_code/exploratory_analysis.qmd
update

## Analysis Data
```{r}
# Path to Data
data_location <- here(
  "data","processed_data","processed_analysis_data.rds")
# Import Data
analysis_data_final <- readRDS(data_location)
```

@@

## Closer Look at Bacterial Etiologies for Outbreaks
ggplot(training_data_final %>% 
         select(Pathogen_Type, Simplified_Etiology, Hospitalizations) %>% filter(Pathogen_Type == "Bacteria"), aes(x = Simplified_Etiology)) + geom_bar() + 
  labs(x = "Bacteria", y = "Hospitalizations", title = "Bar Chart of Bacteria Frequency")+
  coord_flip()

@@

####

Norovirus_Hosp=training_data_final%>%
  select(Simplified_Etiology, Hospitalizations, `IFSAC Category`)%>%
  group_by(Simplified_Etiology, Hospitalizations, `IFSAC Category`)%>%
  filter(Simplified_Etiology=="Salmonella")%>%
  filter(`IFSAC Category` != "Multiple")%>%
  filter(`IFSAC Category` != "Other") %>%
  count()%>%
  mutate(total_hospitalizations=n*Hospitalizations)

ggplot(Norovirus_Hosp,aes(x=`IFSAC Category`,y=total_hospitalizations))+
  geom_boxplot()+
  coord_flip()

@@

## Project total hospitalizations in states map
training_hosp=training_data_final%>%
  mutate(state=tolower(State))%>%
  group_by(state)%>%
  mutate(all_hosp = sum(Hospitalizations)) %>%  count(all_hosp) %>%
  select(state, all_hosp)
training_hosp

state_map=training_hosp%>%
  distinct(state)%>%
  inner_join(.,map_data("state"),by=c("state"="region"))
state_map

map_final=right_join(training_hosp,state_map,by=c("state"="state"))

map_final%>%
  ggplot(aes(x=long,y=lat,fill=all_hosp,group=group))+
  geom_polygon(color="black",size=0.1)+
  geom_polygon(data=state_map,aes(x=long,y=lat,group=group),fill=NA,color="black",size=0.3)+
  scale_fill_gradient(name="Number of \n Hospitalizations",
                      low="#FFFFFF",high="#FF0000",
                       limits=c(0,NA))+
  labs(title="Cases of Hospitalizations Caused by Foodborne Illnesses\n in Each State During 1998-2015")+
  theme_void()+
  coord_quickmap()

ggsave("State_hospitalization_map.png",path=image_path,width=3000,height=2000,units="px")

@@

```
All outbreaks with a high hospitalization count (> 200 cases) are caused by bacteria. It also appears that most outbreaks caused by bacteria are due to Salmonella. California and Minnesota are the two states with the highest total hospitalization count due to foodborne illness from 1998 to 2015. Restaurants and Private home are the most frequent locations with the highest total hospitalization count.

>>>>

## Testing/Analysis Data
```{r}
# Path to Data
data_location <- here(
  "data","processed_data","processed_analysis_data.rds")
# Import Data
analysis_data_final <- readRDS(data_location)
```

@@

## Closer Look at Bacterial Etiologies for Outbreaks
ggplot(training_data_final %>% 
         select(Pathogen_Type, Simplified_Etiology, Hospitalizations) %>% filter(Pathogen_Type == "Bacteria"), aes(x = Simplified_Etiology)) + geom_bar() + 
  labs(x = "Bacteria", y = "Hospitalizations", 
       title = "Bar Chart of Bacteria Frequency") +
  coord_flip()

@@

### Even Closer Look at Salmonella Cases

Salmonella_Hosp <- training_data_final %>%
  select(Simplified_Etiology, Hospitalizations, `IFSAC Category`) %>%
  group_by(Simplified_Etiology, Hospitalizations, `IFSAC Category`) %>%
  filter(Simplified_Etiology == "Salmonella") %>%
  filter(`IFSAC Category` != "Multiple") %>%
  filter(`IFSAC Category` != "Other") %>%
  count() %>%
  mutate(total_hospitalizations = n * Hospitalizations) %>%
  arrange(desc(total_hospitalizations))

ggplot(Salmonella_Hosp, aes(x=`IFSAC Category`, y = total_hospitalizations))+
  geom_boxplot() +
  coord_flip()

@@

## Project Total Hospitalizations onto State Map
training_hosp <- training_data_final%>%
  mutate(state = tolower(State))%>%
  group_by(state)%>%
  mutate(all_hosp = sum(Hospitalizations)) %>% count(all_hosp) %>%
  select(state, all_hosp)
training_hosp

state_map <- training_hosp %>%
  distinct(state)%>%
  inner_join(., map_data("state"), by = c("state"="region"))

state_map

map_final <- right_join(training_hosp,state_map,by=c("state"="state"))

map_final %>%
  ggplot(aes(x=long, y=lat, 
             fill = all_hosp, group = group)) +
  geom_polygon(color="black", size = 0.1) +
  geom_polygon(data = state_map, aes(x= long, y= lat, group = group), 
               fill = NA, color = "black", size= 0.3) +
  scale_fill_gradient(name = "Number of \n Hospitalizations",
                      low = "#FFFFFF", high="#FF0000",
                       limits = c(0,NA)) +
  labs(title = "Cases of Hospitalizations Caused by Foodborne Illnesses\n in Each State During 1998-2015") +
  theme_void() +
  coord_quickmap()

ggsave("State_hospitalization_map.png", path = image_path, width=3000, height=2000, units="px")

@@

```

All outbreaks with a high hospitalization count (> 200 cases) are caused by bacteria. It also appears that most outbreaks caused by bacteria are due to Salmonella. California and Minnesota are the two states with the highest total hospitalization count due to foodborne illness from 1998 to 2015. Restaurants and Private home are the most frequent locations with the highest total hospitalization count.

### 2023-04-07 code/analysis_code/statistical_analysis.qmd
update

# Load libraries
```{r}
# Pathing
library(here)

# Data Handling and Modeling
library(tidyverse)
library(tidymodels)
library(poissonreg)
```
# Load Data
```{r}
# Path to Data
data_location <- here(
  "data","processed_data","processed_training_data.rds")
# Import Data
training_data_final <- readRDS(data_location)
glimpse(training_data_final)
```
# First Health Outcome of Interest: Illnesses

Predictors
-   Time (Year)
-   Pathogen Type
-   Etiology
-   Location
-   State
-   Food Type (IFSAC Category)

```{r}
# Checking Illness Distribution
training_data_final %>%
  ggplot(aes(Illnesses))+
  geom_histogram()+
  scale_x_continuous(trans=scales::pseudo_log_trans(base=10))
image_path <- here("results")
ggsave("histogram_Illnesses.png", path = image_path, width = 3000, height = 2000, units = "px")

# Tibble with Predictors of Interest
illness_data=training_data_final %>%
  select(c(1, 4, 5, 7, 10, 13, 14)) %>%
  mutate_if(is.character, as.factor)
glimpse(illness_data)
```

## Modeling Illness with Linear Regression and Poisson Regression

@@

```{r}
# Linear Regression
lm_spec = linear_reg() %>%
  set_engine(engine = "lm")
## Modeling with Main Predictor `Simplified_Etiology`
lm_fit1 = lm_spec %>%
  fit(Illnesses ~ Simplified_Etiology, data = illness_data)
## Modeling with All Predictors
lm_fit2 = lm_spec %>%
  fit(Illnesses ~ ., data = illness_data)

# Poisson Regression
## Modeling with Main Predictor `Simplified_Etiology`
poisson_wf1 = workflow(Illnesses ~ Simplified_Etiology, poisson_reg())
## Modeling with All Predictors
poisson_wf2 = workflow(Illnesses ~ ., poisson_reg())
## Model Fitting
poisson_fit1 = fit(poisson_wf1, data=illness_data)
poisson_fit2 = fit(poisson_wf2, data=illness_data)
```

### Model Evaluation
```{r}
# Checking Parameters for Linear Regression Model
tidy(lm_fit1) %>%
  arrange(-estimate)
tidy(lm_fit2) %>%
  arrange(-estimate)

# Checking Parameters for Poisson Regression Model
tidy(poisson_fit1) %>%
 arrange(-estimate)
tidy(poisson_fit2) %>%
 arrange(-estimate)

# Model Evaluation
Result = lm_fit1 %>% 
  predict(new_data = illness_data) %>% 
  mutate(.pred = round(.pred), 
        true_value = illness_data$Illnesses, model= "lm_single") %>% 
  bind_rows(lm_fit1 %>% 
              predict(new_data=illness_data) %>% 
              mutate(.pred=round(.pred), 
                     true_value = illness_data$Illnesses, model= "lm_multiple") %>% 
              bind_rows(poisson_fit1 %>% predict(new_data= illness_data) %>%
                          mutate(true_value=illness_data$Illnesses, 
                                 model= "glm_single") %>% 
                          bind_rows(poisson_fit2 %>% 
                                      predict(new_data= illness_data)%>% 
                                      mutate(true_value=illness_data$Illnesses, 
                                             model= "glm_multiple")))
)

Result <- Result %>%
  group_by(model) %>%
  rmse(truth=true_value, estimate = .pred) %>%
  rename(illness_model = model)

saveRDS(Result, file = here("results", "illness_model_evaluation.rds"))

```
According to the result, the model built using Poisson regression with multiple predictors seems to have the smallest root mean squared error.

# Second Health Outcome of Interest: Hospitilization

Predictors
-   Time (Year)
-   Pathogen Type
-   Etiology
-   Location
-   State
-   Food Type (IFSAC Category)

```{r}
# Checking the Distribution of Hospitalization
training_data_final %>%
  ggplot(aes(Hospitalizations)) +
  geom_histogram() +
  scale_x_continuous(trans=scales::pseudo_log_trans(base = 10))
# Tibble for Variables of Interest
hosp_data = training_data_final %>%
  select(c(1, 4, 5, 7, 11, 13, 14))%>%
  mutate_if(is.character, as.factor)
glimpse(hosp_data)
ggsave("histogram_Hospitalizations.png", path = image_path)
```

## Modeling `Hospitalizations` using Linear Regression and Poisson Regression

@@

```{r}
# Model Specification and Fitting for Linear Regression Model
lm_spec = linear_reg() %>%
  set_engine(engine = "lm")
## Modeling with Main Predictor `Simplified_Etiology`
lm_fit3 = lm_spec %>%
  fit(Hospitalizations ~ Simplified_Etiology, data = hosp_data)
## Modeling with All Predictors
lm_fit4 = lm_spec %>%
  fit(Hospitalizations ~ ., data = hosp_data)

# Model Specification and Fitting for Poisson Regression Model
## Modeling with Main Predictor `Simplified_Etiology`
poisson_wf3 = workflow(Hospitalizations ~ Simplified_Etiology, poisson_reg())
## Modeling with all predictors
poisson_wf4=workflow(Hospitalizations~., poisson_reg())
## Fitting models
poisson_fit3 = fit(poisson_wf3,data = hosp_data)
poisson_fit4 = fit(poisson_wf4,data = hosp_data)
```

### Model Evaluation
```{r}
Result = lm_fit3 %>%
 predict(new_data = hosp_data) %>%
 mutate(.pred = round(.pred), true_value = hosp_data$Hospitalizations,
        model= "lm_single") %>% 
  bind_rows(lm_fit4 %>% 
              predict(new_data=hosp_data) %>% 
              mutate (.pred=round(.pred), 
                      true_value=hosp_data$Hospitalizations, 
                      model= "lm_multiple") %>% 
              bind_rows(poisson_fit1 %>% 
                        predict(new_data = hosp_data) %>%
                        mutate(true_value = hosp_data$Hospitalizations, 
                               model= "glm_single") %>% 
                          bind_rows(poisson_fit2 %>% 
                                      predict(new_data= hosp_data) %>% 
                                      mutate(true_value = hosp_data$Hospitalizations, 
                                             model= "glm_multiple")))
)

Result <- Result %>%
  group_by(model)%>%
  rmse(truth = true_value, estimate = .pred) %>%
  rename(hospital_model = model)

saveRDS(Result, file = here("results", "hospital_model_evaluation.rds"))

@@

According to the result, the model built using linear regression with multiple predictors seems to have the smallest root mean squared error.

>>>>

---
title: "Statistical analysis script"
author: "Weifan Wu and Kailin Chen"
date: "03/16/2023"
output: html_document
---

# Load Libraries
```{r}
# Pathing
library(here)

# Data Handling and Modeling
library(tidyverse)
library(tidymodels)
library(poissonreg)
library(knitr)
library(skimr)
library(vip)
## For Decision Trees
library(ranger)
library(caret)
library(rpart)
library(ipred)
library(e1071)
```
# Load Data
```{r}
# Import Training Data
data_location <- here(
  "data","processed_data","processed_training_data.rds")
training_data_final <- readRDS(data_location)
skim(training_data_final)

# Import Test Data
data_location <- here(
  "data","processed_data","processed_testing_data.rds")
test_data_final <- readRDS(data_location)
skim(test_data_final)

```{r}
# Set Image Path
images_stats_path <- here("results","images_statistic_analysis")

# Checking General Illness Distribution
training_data_final %>%
  ggplot(aes(Illnesses)) +
  geom_histogram() +
  scale_x_continuous(trans = scales::pseudo_log_trans(base = 10))
## Save Histogram
ggsave("histogram_Illnesses.png", path = images_stats_path, width = 3000, height = 2000, units = "px")

# Checking General Hospitalization Distribution
training_data_final %>%
  ggplot(aes(Hospitalizations)) +
  geom_histogram() +
  scale_x_continuous(trans=scales::pseudo_log_trans(base = 10))
## Save Generated Histogram
ggsave("histogram_Hospitalizations.png", path = images_stats_path)
```

# Create Tibbles for Illness & Hospitalizations
```{r}
process_datasets_to_tibbles <- function(dataset) {
  dataset %>%
    filter(!`IFSAC Category` %in% c("Multiple", "Other"), 
           Location!= "Unknown") %>%
    mutate_if(is.character, as.factor) %>%
    select(!Pathogen_Type)
}

illness_data <- process_datasets_to_tibbles(training_data_final) %>%
  select(!Month) %>%
  rename(Month = Numeric_Month) %>%
  select(!Food) %>% select(!Hospitalizations)

hosp_data <- process_datasets_to_tibbles(training_data_final) %>%
  select(!Month) %>%
  rename(Month = Numeric_Month) %>%
  select(!Food) %>% select(!Illnesses)

skim(illness_data)
skim(hosp_data)

test_data_final <- process_datasets_to_tibbles(test_data_final) %>%
  select(!Food)

skim(test_data_final)
```

The two health outcomes of interest are illnesses and hospitalizations.

Predictors
-   Time (Month, Year)
-   Pathogen Type
-   Etiology
-   Location
-   State
-   Food Type (IFSAC Category)

# Modeling Illness

@@

```{r}
# Set Seed for Reproducibility
set.seed(123)

# Cross-Validation Setup
illness_fold <- vfold_cv(illness_data, repeats = 2, strata = Illnesses)

# Additional Preprocessing to Deal with Variables with > 50 Categories
preprocessing_recipe <- function(anterior) {
  anterior %>%
    step_dummy(Simplified_Etiology) %>%
    step_dummy(State) %>%
    step_dummy(`IFSAC Category`)
}

## Simple Recipe for Illnesses
illness_simple_recipe <- recipe(Illnesses ~ Simplified_Etiology, 
                                data = illness_data) %>%
  step_dummy(Simplified_Etiology)

## Recipe with All Predictors for Illnesses
illness_complex_recipe <- recipe(Illnesses ~ ., data = illness_data) %>%
  preprocessing_recipe()

### Lump small proportion into "other" using step_other
### Create dummy code for nominal variables
#illness_all_rec=recipe(Illnesses~.,data=illness_data)%>%
#  step_other(State,Location,`IFSAC Category`,Simplified_Etiology,threshold = 0.01)%>%
#  step_dummy(all_nominal())

### Check Processed Dataset
illness_complex_recipe %>% prep() %>% bake(new_data = NULL)

# Model Specification
## Linear Regression
linear_model <- linear_reg() %>%
  set_engine(engine = "lm")

# Create Workflow
## Main Predictor `Simplified_Etiology`
linear_workflow_simple_illness <- workflow() %>%
  add_recipe(illness_simple_recipe) %>%
  add_model(linear_model)

## All Predictors
linear_workflow_complex_illness <- workflow() %>%
  add_recipe(illness_complex_recipe)%>%
  add_model(linear_model)

## Poisson Regression
## Modeling with Main Predictor `Simplified_Etiology`
poisson_workflow_simple_illness <- workflow() %>%
  add_recipe(illness_simple_recipe) %>%
  add_model(poisson_reg())

## Modeling with All Predictors
poisson_workflow_complex_illness <- workflow() %>%
  add_recipe(illness_complex_recipe) %>%
  add_model(poisson_reg())
```

## Linear Regression
### Model Fitting
```{r}
set.seed(234)

## Linear Regression
### Simple Model
lm_simple_resampled_illness <- linear_workflow_simple_illness %>%
  fit_resamples(resamples = illness_fold, control = control_resamples(save_pred = TRUE))

### All Predictors
lm_complex_resampled_illness <- linear_workflow_complex_illness %>%
  fit_resamples(resamples = illness_fold, control = control_resamples(save_pred = TRUE))
```

### Model Evaluation
```{r}
# Check Metrics
## Simple Regression
lm_simple_resampled_illness %>%
  collect_metrics()
## Multivariate Regression
lm_complex_resampled_illness %>%
  collect_metrics()

# Check Parameters for Linear Regression Model
## Simple Regression
linear_model_simple_fit_illness <- linear_workflow_simple_illness %>%
  fit(illness_data) %>%
  tidy() %>%
  arrange(-estimate)
## Multivariate Regression
linear_model_complex_fit_illness <- linear_workflow_complex_illness %>%
  fit(illness_data) %>%
  tidy() %>%
  arrange(-estimate)

# Check Variable Importance
## Simple Regression
linear_workflow_simple_illness %>%
  fit(illness_data) %>%
  extract_fit_engine() %>%
  vip()
## Multivariate Regression
linear_workflow_complex_illness %>%
  fit(illness_data) %>%
  extract_fit_engine() %>%
  vip()

# Plotting Predicted Values vs True Values (Illnesses)
lm_simple_resampled_illness %>%
  collect_predictions() %>%
  # filter(Illnesses < 300) %>%
  ggplot(aes(Illnesses, .pred, color = id2, group = id2)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  geom_abline(intercept = 0,slope = 1) +
  labs(y = "Predicted Illnesses", 
       title = "Predicted Illnesses VS Illnesses via Simple Linear Regression")

ggsave("Predicted Illnesses VS Illnesses by linear regression with main predictor.png", path = images_stats_path, width = 3000, height = 2000, units = "px")

lm_complex_resampled_illness %>%
  collect_predictions() %>%
  # filter(Illnesses < 300) %>%
  ggplot(aes(Illnesses, .pred, color = id2, group = id2)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  geom_abline(intercept = 0, slope = 1) +
  labs(y = "Predicted Illnesses", title = "Predicted Illnesses VS Illnesses via Multivariate Linear Regression")

ggsave("Predicted Illnesses VS Illnesses by linear regression with all predictors.png", path=images_stats_path, width = 3000, height = 2000, units = "px")

## Poisson Regression
### Model Fitting
```{r}
## Poisson Regression
### Simple
poisson_simple_resampled_illness <- poisson_workflow_simple_illness %>%
  fit_resamples(resamples = illness_fold,
    control = control_resamples(save_pred = TRUE))

### Multivariate
poisson_complex_resampled_illness <- poisson_workflow_complex_illness %>%
  fit_resamples(resamples = illness_fold,
    control = control_resamples(save_pred = TRUE))
```

### Model Evaluation
```{r}
# Metrics
## Simple
poisson_simple_resampled_illness %>%
  collect_metrics()
## Complex
poisson_complex_resampled_illness %>%
  collect_metrics()

# Check Parameters
## Simple
poisson_simple_fit_illness <- poisson_workflow_simple_illness %>%
  fit(illness_data) %>%
  tidy() %>%
  arrange(-estimate)
## Complex
poisson_complex_fit_illness <- poisson_workflow_complex_illness %>%
  fit(illness_data) %>%
  tidy() %>%
  arrange(-estimate)

# Check Variable Importance
## Simple
poisson_workflow_simple_illness %>%
  fit(illness_data) %>%
  extract_fit_engine() %>%
  vip()
## Complex
poisson_workflow_complex_illness %>%
  fit(illness_data) %>%
  extract_fit_engine() %>%
  vip()

# Plotting Predictions vs Illnesses
## Simple
poisson_simple_resampled_illness %>%
  collect_predictions() %>%
  # filter(Illnesses < 300) %>%
  ggplot(aes(Illnesses, .pred, color = id2, group = id2)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  geom_abline(intercept = 0, slope = 1) +
  labs(y = "Predicted Illnesses", title = "Predicted Illnesses VS Illnesses by Simple Poisson Regression")

ggsave("Predicted Illnesses VS Illnesses by poisson regression with main predictor.png", path = images_stats_path, width = 3000, height = 2000, units = "px")

## Complex
poisson_complex_resampled_illness %>%
  collect_predictions() %>%
  # filter(Illnesses < 500) %>%
  ggplot(aes(Illnesses, .pred, color = id2, group = id2)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  geom_abline(intercept = 0,slope = 1) +
  labs(y = "Predicted Illnesses", title = "Predicted Illnesses VS Illnesses by linear regression with all predictors")

ggsave("Predicted Illnesses VS Illnesses by poisson regression with all predictors.png", path = images_stats_path, width = 3000, height = 2000, units = "px")

```
According to the result, the model built using multivariate linear regression seems to have the smallest root mean squared error and highest R^2. However, none of those models can catch outliers (Illness > 1000).

# Modeling Hospitilizations

@@

```{r}
# Cross-Validation
hosp_fold <- vfold_cv(hosp_data, repeats = 2, strata = Hospitalizations)

# Data Preprocessing
## Simple Regression Recipe
hosp_simple_recipe <- recipe(Hospitalizations ~ Simplified_Etiology, data = hosp_data) %>%
  step_dummy(Simplified_Etiology)

## Multivariate Regression Recipe
hosp_complex_recipe <- recipe(Hospitalizations ~ ., data = hosp_data) %>%
  preprocessing_recipe()

## Recipe for all predictors
### Lump small proportion into "other" using step_other
### Create dummy code for nominal variables
# hosp_all_rec=recipe(Hospitalizations ~.,data=hosp_data)%>%
#  step_other(State,Location,`IFSAC Category`, Simplified_Etiology,threshold = 0.01)%>%
#  step_dummy(all_nominal_predictors())

# Check Preprocessed Dataset
hosp_complex_recipe %>%
  prep() %>%
  bake(new_data = NULL)

# Create Workflows
## Simple
linear_workflow_simple_hosp <- workflow() %>%
  add_recipe(hosp_simple_recipe) %>%
  add_model(linear_model)
poisson_workflow_simple_hosp <- workflow() %>%
  add_recipe(hosp_simple_recipe) %>%
  add_model(poisson_reg())
## Multivariate
linear_workflow_complex_hosp <- workflow() %>%
  add_recipe(hosp_complex_recipe) %>%
  add_model(linear_model)
poisson_workflow_complex_hosp <- workflow() %>%
  add_recipe(hosp_complex_recipe) %>%
  add_model(poisson_reg())
```

## Linear Regression
```{r}
set.seed(234)
```

### Model Fitting
```{r}
# Simple
lm_simple_resampled_hosp <- linear_workflow_simple_hosp %>%
  fit_resamples(resamples = hosp_fold,
    control = control_resamples(save_pred = TRUE))

# Complex
lm_complex_resampled_hosp <- linear_workflow_complex_hosp %>%
  fit_resamples(resamples = hosp_fold,
    control = control_resamples(save_pred = TRUE))
```

### Model Evaluation
```{r}
# Metrics
lm_simple_resampled_hosp %>%
  collect_metrics()
lm_complex_resampled_hosp %>%
  collect_metrics()

# Check Parameters
linear_workflow_simple_hosp %>%
  fit(hosp_data) %>%
  tidy() %>%
  arrange(-estimate)
linear_workflow_complex_hosp %>%
  fit(hosp_data) %>%
  tidy() %>%
  arrange(-estimate)

## Visualizing Variable Importance
linear_workflow_simple_hosp %>%
  fit(hosp_data) %>%
  extract_fit_engine() %>%
  vip()
linear_workflow_complex_hosp %>%
  fit(hosp_data) %>%
  extract_fit_engine() %>%
  vip()

# Plotting predictions vs Hospitalizations
lm_simple_resampled_hosp %>%
  collect_predictions() %>%
  # filter(Hospitalizations < 50) %>%
  ggplot(aes(Hospitalizations, .pred, color = id2))+
  geom_point() +
  geom_smooth(se = FALSE) +
  geom_abline(intercept = 0, slope = 1) +
  labs(y = "Predicted Hospitalizations", 
       title = "Predicted Hospitalizations VS Hospitalizations \by Simple Linear Regression", color = "Fold")

ggsave("Predicted Hospitalizations VS Hospitalizations by linear regression with main predictor.png", path = images_stats_path, width = 3000, height = 2000, units = "px")

lm_complex_resampled_hosp %>%
  collect_predictions() %>%
  #filter(Hospitalizations < 50) %>%
  ggplot(aes(Hospitalizations, .pred, color = id2))+
  geom_point() +
  geom_smooth(se = FALSE) +
  geom_abline(intercept = 0, slope = 1) +
  labs(y = "Predicted Hospitalizations", 
       title="Predicted Hospitalizations VS Hospitalizations \by Multivariate Linear Regression", color = "Fold")

ggsave("Predicted Hospitalizations VS Hospitalizations by linear regression with all predictors.png", path = images_stats_path, width = 3000, height = 2000, units = "px")
```
## Poisson Regression
### Model Fitting
```{r}
## Poisson Regression
### Simple
poisson_simple_resampled_hosp <- poisson_workflow_simple_hosp %>%
  fit_resamples(resamples = hosp_fold,
    control = control_resamples(save_pred = TRUE))

### Multivariate
poisson_complex_resampled_hosp <- poisson_workflow_complex_hosp %>%
  fit_resamples(resamples = hosp_fold,
    control = control_resamples(save_pred = TRUE))
```

### Model Evaluation
```{r}
# Metrics
poisson_simple_resampled_hosp %>%
  collect_metrics()
poisson_complex_resampled_hosp %>%
  collect_metrics()

# Check Parameters
poisson_workflow_simple_hosp %>%
  fit(hosp_data) %>%
  tidy() %>%
  arrange(-estimate)

poisson_workflow_complex_hosp %>%
  fit(hosp_data) %>%
  tidy() %>%
  arrange(-estimate)

# Variable Importance
poisson_workflow_simple_hosp %>%
  fit(hosp_data) %>%
  extract_fit_engine() %>%
  vip()
poisson_workflow_complex_hosp %>%
  fit(hosp_data) %>%
  extract_fit_engine() %>%
  vip()

# Predictions vs Hospitalizations
poisson_simple_resampled_hosp %>%
  collect_predictions() %>%
  # filter(Hospitalizations < 50) %>%
  ggplot(aes(Hospitalizations, .pred, color = id2)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  geom_abline(intercept = 0,slope = 1) +
  labs(y = "Predicted Hospitalizations", 
       title = "Predicted Hospitalizations VS Hospitalizations \by Simple Linear Regression", color = "Fold")

ggsave("Predicted Hospitalizations VS Hospitalizations by poisson regression with main predictor.png", path = images_stats_path, width = 3000, height = 2000, units = "px")

poisson_complex_resampled_hosp %>%
  collect_predictions() %>%
  # filter(Hospitalizations < 50) %>%
  ggplot(aes(Hospitalizations, .pred, color = id2)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  geom_abline(intercept = 0, slope = 1) +
  labs(y = "Predicted Hospitalizations", title = "Predicted Hospitalizations VS Hospitalizations \by Multivariate Linear Regression", color = "Fold")

ggsave("Predicted Hospitalizations VS Hospitalizations by poisson regression with all predictors.png", path = images_stats_path, width = 3000, height = 2000, units = "px")

```
According to the results, the model built using multivariate linear regression seems to have the smallest root mean squared error and highest R^2 (14%). But again, outliers (Hospitalizations > 100) cannot be predicted well by any of those models.

# Machine Learning Approach
```{r}
set.seed(345)
```

## Random Forest
### Predicting First Outcome `Illnesses` Using All Predictors
```{r}
# Model Specification
rf_spec <- rand_forest(mtry = tune(), min_n = tune(), trees = 1000) %>%
  set_engine("ranger", importance = 'impurity') %>%
  set_mode("regression")

# illness_all_rec <- recipe(Illnesses~.,data=illness_data) %>%
#  step_other(State,Location,`IFSAC Category`,Simplified_Etiology,threshold = 0.01)%>%
#  step_dummy(all_nominal())

# Workflow
random_forest_illness_workflow <- workflow() %>%
  add_recipe(illness_complex_recipe) %>%
  add_model(rf_spec)

# Training and Tuning Model
doParallel::registerDoParallel()
random_forest_illness_tuned <- random_forest_illness_workflow %>%
  tune_grid(
    resamples = illness_fold,
    grid = 15,
    control = control_grid(save_pred = TRUE),
    metrics = metric_set(rmse)
  )
```

### Model Evaluation
```{r}
# Plot
random_forest_illness_tuned %>%
  autoplot()

# Final Specifications
random_forest_illness_tuned_best <- random_forest_illness_tuned %>%
  select_best()
random_forest_illness_final <- random_forest_illness_workflow %>%
  finalize_workflow(random_forest_illness_tuned_best)
random_forest_illness_fit <- random_forest_illness_final %>%
  fit(illness_data)

# Plotting Predicted Illness VS Illnesses
random_forest_illness_predict <- random_forest_illness_fit %>%
  predict(illness_data)

illness_data %>%
  select(Illnesses) %>%
  bind_cols(random_forest_illness_predict) %>%
  ggplot(aes(Illnesses, .pred)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1)+
  geom_smooth(se = FALSE) +
  labs(y = "Predicted Illnesses", 
       title = "Predicted Illnesses VS Illnesses by Random Forest Model")
ggsave("Predicted illnesses VS illnesses by random forest model.png", path = images_stats_path, width = 3000, height = 2000, units = "px")
```

### Predicting Second Outcome `Hospitalizations` Using All Predictors
```{r}
set.seed(456)

# Creating workflow
random_forest_hosp_workflow <- workflow() %>%
  add_recipe(hosp_complex_recipe) %>%
  add_model(rf_spec)
# Training and tuning model
doParallel::registerDoParallel()
random_forest_hosp_tuned <- random_forest_hosp_workflow %>%
  tune_grid(
    resamples = hosp_fold,
    grid = 15,
    control = control_grid(save_pred = TRUE),
    metrics = metric_set(rmse)
  )
```

### Model Evaluation
```{r}
# Plot
random_forest_hosp_tuned %>%
  autoplot()

# Final Specifications
random_forest_hosp_tuned %>%
  show_best()
random_forest_hosp_best <- random_forest_hosp_tuned %>%
  select_best()

random_forest_hosp_final <- random_forest_hosp_workflow %>%
  finalize_workflow(random_forest_hosp_best)

random_forest_hosp_fit <- random_forest_hosp_final %>%
  fit(hosp_data)

random_forest_hosp_predict <- random_forest_hosp_fit %>%
  predict(hosp_data)

hosp_data %>%
  select(Hospitalizations) %>%
  bind_cols(random_forest_hosp_predict) %>%
  ggplot(aes(Hospitalizations, .pred)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1) +
  geom_smooth(se = FALSE) +
  labs(y = "Predicted Hospitalizations",
       title = "Predicted Hospitalizations VS Hospitalizations by Random Forest")

ggsave("Predicted Hospitalizations VS Hospitalizations by random forest model.png", path = images_stats_path, width = 3000, height = 2000, units = "px")
```

Overall, random forest model has the lowest rmse among those selected models when training data is fitted; Also, the random forest model outperforms other models in predicting outliers.

## Bagging
### Illnesses
```{r}
bag_training_illness <- bagging(formula = Illnesses ~ .,
                        data = illness_data,
                        method = "treebag",
                        trControl = trainControl(method = "cv", number = 30),
                        nbagg = 500,
                        coob = TRUE,
                        control = rpart.control(minsplit = 2, cp = 0))

bag_training_illness
```

### Hospitalizations
```{r}
bag_training_hosp <- bagging(formula = Hospitalizations ~ .,
                        data = hosp_data,
                        method = "treebag",
                        trControl = trainControl(method = "cv", number = 30),
                        nbagg = 500,
                        coob = TRUE,
                        control = rpart.control(minsplit = 2, cp = 0))

bag_training_hosp
```

## Organize Model Performance and Choose the Best One
### Illnesses
```{r}
# Linear Regression
lm_complex_resampled_illness %>%
  collect_metrics()

# Poisson Regression
poisson_complex_resampled_illness %>%
  collect_metrics()

# Random Forest
random_forest_illness_predict %>%
  collect_metrics()

# Make Dataframe of Model RMSEs

@@

### Hospitalizations
```{r}
# Linear Regression
# Poisson Regression
# Random Forest

## Fit Testing Data
